### 2020-01-28 thesis/chapter-1.tex
update thesis bib and introduction.

% \begin{enumerate}
%     \item Chapter 1 – General Introduction Links all manuscripts
%  \item Chapter 2 – First Manuscript
%  \item Chapter 3 – Second Manuscript
%  \item Chapter 4 – Third Manuscript
%  \item Chapter 6 – General Conclusion (common conclusion linking all manuscripts
% thematically)
%  \item Common bibliography covering all manuscripts, although each manuscript may have its own reference section
% \end{enumerate}

@@

\subsection{Definition of Entropy $S(E)$}\label{entropy}
The entropy $S(E)$ can be defined for any given system $\Gamma$ with discrete energy states using one of the most important formulae in physics:
\begin{align}\label{sdos}
    S(E,V,N) \equiv k_B \ln\Omega(E,V,N)
\end{align}
where the entropy is defined as the logarithm of the density of states and $k_B$ is the Boltzmann constant. An important aspect of this definition in terms of logarithms is that the entropy for multiple isolated systems is additive.
\begin{align}
    S(E_1,E_2) &= k_B \ln(\Omega(E_1)\Omega_2(E_2)) \\
    &= k_B \left( \Omega_1(E_1) + \Omega_2(E_2)\right) \\
    &= k_B \left( S_1(E_1) + S_2(E_2)\right)
\end{align}
From the \nth{1} law of thermodynamics, the system temperature can be written in terms of the density of states by refering to Eq.~\HR{sdos}.
\begin{align}
    T = \left(\diff{S(E)}{E}\right)_V = \frac{k_B}{\Omega(E)}\left(\diff{\Omega(E)}{E}\right)_V
\end{align}
The relationship between entropy $S(E)$, $\Omega(E)$, and $T$ will be discussed in further depth in sections \HR{laws} and \HR{weights}.

@@

\subsection{Maxwell Relations}\label{maxwell}
Maxwell's relations are second derivative equalities that relate important thermodynamic variables. They are presented here to facilitate a critical derivation in Section~\HR{gas-ads}. Maxwell relations are an extension of the Schwarz theorem which states that the order of differentiation is irrelevant.
\begin{align}
    \pdv{}{\theta}\left(\pdv{\Theta}{\phi}\right)=\pdv{}{\phi}\left(\pdv{\Theta}{\theta}\right)
\end{align}
From this theorem and Section~\HR{potentials}, we can derive the most common of Maxwell's relations. The relation involving internal energy $U\left(S,V,N\right)$ relates the natural variables of pressure $p$ and temperature $T$ and can be derived using Equation~\ref{maxU}.
\begin{align}
    \left( {\pdv{U}{S}{V}} \right )_{N} = \left( \pdv{T}{V}\right )_{S,N} = - \left( \pdv{p}{S}\right )_{V,N}
\end{align}
The relation involving the Helmholtz free energy $F\left(T,V,N\right)$ relates the natural variables of entropy $S$ and pressure $p$ and can be derived using Equation~\ref{maxF}.
\begin{align}
    \left( {\pdv{F}{T}{V}} \right )_{N} = \left( \pdv{S}{V}\right )_{T,N} = \left( \pdv{p}{T}\right )_{V,N}
\end{align}
The relation involving the Enthalpy $H\left(S,p,N\right)$ relates the natural variables of temperature $T$ and volume $V$ and can be derived using Equation~\ref{maxH}.
\begin{align}
    \left( {\pdv{H}{S}{p}} \right )_{N} = \left( \pdv{T}{p}\right )_{S,N} =  \left( \pdv{V}{S}\right )_{p,N}
\end{align}
The relation involving the Gibbs free energy $G\left(T,p,N\right)$ relates the natural variables of entropy $S$ and volume $V$ and can be derived using Equation~\ref{maxG}.
\begin{align}
    \left( {\pdv{G}{T}{p}} \right )_{N} = -\left( \pdv{S}{p}\right )_{T,N} = - \left( \pdv{V}{T}\right )_{p,N}
\end{align}

\subsection{Thermodynamic Properties}\label{thermoprop}
The Maxwell relations from Section~\HR{maxwell} can be written in terms of thermodynamic properties of the material.  The isochoric heat capacity $C_V$ of a given material is defined as the amount of heat necessary to raise the temperature by a given amount.
\begin{align}
    C_V = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{V,N} = \frac{1}{T}\left(\pdv{S}{T}\right )_{V,N}
\end{align}
The isobaric heat capacity $C_p$, which is always greater than or equal to $C_V$ according to the Mayer relation, can also be written in terms of the entropy $S$ and temperature $T$ of the system.
\begin{align}
    C_p = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{p,N} = T\left(\pdv{S}{T}\right )_{p,N}
\end{align}
The Mayer relation gives the difference of the isobaric and isochoric heat capacities in terms of the thermal expansion coefficient and the isothermal compressibility.
\begin{align}
    C_p - C_V = VT\frac{\alpha^2}{\kappa_T}
\end{align}
The isobaric coefficient of thermal expansion determines to what extent a given material will deform as a result of a change in temperature.
\begin{align}
    \alpha_p = \frac{1}{V}\left( \pdv{V}{T} \right )_{p,N}
\end{align}
Similarly, the isochoric coefficient of thermal expansion can be written as the following:
\begin{align}
    \alpha_V = \frac{1}{p}\left( \pdv{p}{T} \right )_{V,N}
\end{align}
The compressibility determines to what extent a given material will change volume based on the pressure applied to the system.  The isothermal compressibility can be written in terms of the following:
\begin{align}
    \kappa_T = -\frac{1}{V}\left( \pdv{V}{p} \right )_{T,N}
\end{align}
Similarly, the isentropic compressibility can be written as the following:
\begin{align}
    \kappa_S = -\frac{1}{V}\left( \pdv{V}{p} \right )_{S,N}
\end{align}
A major driving force behind introducing the concepts of thermal expansion and compressibility is to right complex derivatives in terms of these parameters.  This will be invaluable as will be shown in Section~\HR{gas-ads}. The internal energy derivative and entropy derivative can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{U}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} - \alpha TV \quad\quad
    \left( \pdv{S}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha T} + \alpha V
\end{align}
Additionally, the enthalpy derivatives can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{H}{T} \right )_{V,N} = C_p + \frac{\alpha V}{\kappa_T}\left(1-\alpha T\right) \quad\quad
    \left( \pdv{H}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} + V\left(1-\alpha T\right)
\end{align}
Finally, the Gibbs free energy derivatives can be written in terms of the isothermal compressibility, the thermal coefficient of expansion, and the entropy.
\begin{align}
    \left( \pdv{G}{T} \right )_{V,N} = \frac{\alpha V}{\kappa_T}-S \quad\quad
    \left( \pdv{G}{p} \right )_{V,N} = V-\frac{\kappa_T S}{\alpha}
\end{align}

@@

\section{Monte Carlo Simulation}\label{monte}

% talk about Canonical here
\subsection{Broad Histogram Methods}\label{bhm}
\subsection{Weights, multiplicity, detailed balance}\label{weights}

\subsection{System Details}
% Ising, Square well, Lennard-Jones clusters, MOF?

\section{Real System Simulation}
% brief outline for broad histogram methods and how the motivate study of MOFs

\subsection{Metal-organic Frameworks (MOFs)}
\subsection{Gas Adsorption}\label{gas-ads}
% qst

@@

% Tie everything in together here!

>>>>

Monte Carlo methods are a powerful simulation tool that allow scientists to explore properties of systems that may otherwise by inaccessible via experimental methods. Monte Carlo methods can be used to solve a variety of physical problems yielding insight into statistical averages due to random processes. During the last two decades considerable research has been done on developing and refining flat-histogram Monte Carlo methods.  These Monte Carlo (MC) methods are capable of exploring all of the energy phase space for all given temperatures in a \emph{single} simulation. In this work, we first outline the necessary statistical mechanics background. We provide a description of MC methods and flat-histogram methods, as well as, our newly developed method SAD (stochastic approximation Monte Carlo with a dynamic update factor). Finally, we discuss physics rich test systems that can be used to explore the convergence and statistical properties of our newly developed method SAD. In the following chapters, we follow the manuscript outline with each chapter corresponding to our published or submitted research on the SAD method applied to various physical systems.

@@

\subsection{Definition of Entropy $S(E)$}\label{entropy}
The entropy $S(E)$ can be defined for any given system $\Gamma$ with discrete energy states using one of the most important formulae in physics:
\begin{align}\label{sdos}
    S(E,V,N) \equiv k_B \ln\Omega(E,V,N)
\end{align}
where the entropy is defined as the logarithm of the density of states and $k_B$ is the Boltzmann constant. An important aspect of this definition in terms of logarithms is that the entropy for multiple isolated systems is additive.
\begin{align}
    S(E_1,E_2) &= k_B \ln(\Omega(E_1)\Omega_2(E_2)) \\
    &= k_B \left( \Omega_1(E_1) + \Omega_2(E_2)\right) \\
    &= k_B \left( S_1(E_1) + S_2(E_2)\right)
\end{align}
From the \nth{1} law of thermodynamics, the system temperature can be written in terms of the density of states by referring to Eq.~\HR{sdos}.
\begin{align}
    T = \left(\diff{S(E)}{E}\right)_V = \frac{k_B}{\Omega(E)}\left(\diff{\Omega(E)}{E}\right)_V
\end{align}
The relationship between entropy $S(E)$, $\Omega(E)$, and $T$ will be discussed in further depth in Section~\HR{laws}.

@@

\subsection{Maxwell Relations}\label{maxwell}
Maxwell's relations are second derivative equalities that relate important thermodynamic variables. Maxwell relations are an extension of the Schwarz theorem which states that the order of differentiation is irrelevant.
\begin{align}
    \pdv{}{\theta}\left(\pdv{\Theta}{\phi}\right)=\pdv{}{\phi}\left(\pdv{\Theta}{\theta}\right)
\end{align}
From this theorem and Section~\HR{potentials}, we can derive the most common of Maxwell's relations. The relation involving internal energy $U\left(S,V,N\right)$ relates the natural variables of pressure $p$ and temperature $T$ and can be derived using Equation~\ref{maxU}.
\begin{align}
    \left( {\pdv{U}{S}{V}} \right )_{N} = \left( \pdv{T}{V}\right )_{S,N} = - \left( \pdv{p}{S}\right )_{V,N}
\end{align}
The relation involving the Helmholtz free energy $F\left(T,V,N\right)$ relates the natural variables of entropy $S$ and pressure $p$ and can be derived using Equation~\ref{maxF}.
\begin{align}
    \left( {\pdv{F}{T}{V}} \right )_{N} = \left( \pdv{S}{V}\right )_{T,N} = \left( \pdv{p}{T}\right )_{V,N}
\end{align}
The relation involving the Enthalpy $H\left(S,p,N\right)$ relates the natural variables of temperature $T$ and volume $V$ and can be derived using Equation~\ref{maxH}.
\begin{align}
    \left( {\pdv{H}{S}{p}} \right )_{N} = \left( \pdv{T}{p}\right )_{S,N} =  \left( \pdv{V}{S}\right )_{p,N}
\end{align}
The relation involving the Gibbs free energy $G\left(T,p,N\right)$ relates the natural variables of entropy $S$ and volume $V$ and can be derived using Equation~\ref{maxG}.
\begin{align}
    \left( {\pdv{G}{T}{p}} \right )_{N} = -\left( \pdv{S}{p}\right )_{T,N} = - \left( \pdv{V}{T}\right )_{p,N}
\end{align}

\subsection{Thermodynamic Properties}\label{thermoprop}
The Maxwell relations from Section~\HR{maxwell} can be written in terms of thermodynamic properties of the material.  The isochoric heat capacity $C_V$ of a given material is defined as the amount of heat necessary to raise the temperature by a given amount.
\begin{align}
    C_V = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{V,N} = \frac{1}{T}\left(\pdv{S}{T}\right )_{V,N}
\end{align}
The isobaric heat capacity $C_p$, which is always greater than or equal to $C_V$ according to the Mayer relation, can also be written in terms of the entropy $S$ and temperature $T$ of the system.
\begin{align}
    C_p = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{p,N} = T\left(\pdv{S}{T}\right )_{p,N}
\end{align}
The Mayer relation gives the difference of the isobaric and isochoric heat capacities in terms of the thermal expansion coefficient and the isothermal compressibility.
\begin{align}
    C_p - C_V = VT\frac{\alpha^2}{\kappa_T}
\end{align}
The isobaric coefficient of thermal expansion determines to what extent a given material will deform as a result of a change in temperature.
\begin{align}
    \alpha_p = \frac{1}{V}\left( \pdv{V}{T} \right )_{p,N}
\end{align}
Similarly, the isochoric coefficient of thermal expansion can be written as the following:
\begin{align}
    \alpha_V = \frac{1}{p}\left( \pdv{p}{T} \right )_{V,N}
\end{align}
The compressibility determines to what extent a given material will change volume based on the pressure applied to the system.  The isothermal compressibility can be written in terms of the following:
\begin{align}
    \kappa_T = -\frac{1}{V}\left( \pdv{V}{p} \right )_{T,N}
\end{align}
Similarly, the isentropic compressibility can be written as the following:
\begin{align}
    \kappa_S = -\frac{1}{V}\left( \pdv{V}{p} \right )_{S,N}
\end{align}
A major driving force behind introducing the concepts of thermal expansion and compressibility is to right complex derivatives in terms of these parameters. The internal energy derivative and entropy derivative can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{U}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} - \alpha TV \quad\quad
    \left( \pdv{S}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha T} + \alpha V
\end{align}
Additionally, the enthalpy derivatives can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{H}{T} \right )_{V,N} = C_p + \frac{\alpha V}{\kappa_T}\left(1-\alpha T\right) \quad\quad
    \left( \pdv{H}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} + V\left(1-\alpha T\right)
\end{align}
Finally, the Gibbs free energy derivatives can be written in terms of the isothermal compressibility, the thermal coefficient of expansion, and the entropy.
\begin{align}
    \left( \pdv{G}{T} \right )_{V,N} = \frac{\alpha V}{\kappa_T}-S \quad\quad
    \left( \pdv{G}{p} \right )_{V,N} = V-\frac{\kappa_T S}{\alpha}
\end{align}

@@

\section{Monte Carlo Simulation}\label{monte}
Experiment and simulation blend together to help form a more complete understanding of our knowledge of how the universe works. Computational `experiments' such as Monte Carlo methods aid in this understanding by allowing scientists to explore system configurations that are not easily accessible to traditional experiment.

In the field of statistical physics, Monte Carlo methods are used to determine equilibrium 
properties (thermal averages) of multi-particle systems.  A number of Monte Carlo simulations 
have been developed over the last century. Some of these methods are completely unique while 
others are more sophisticated variants of an original method. All of the methods share a number 
of similarities including picking random moves (we will discuss this process in Section~\HR{rng}) and determining the acceptance or rejection of a 
move based on a probability distribution. They differ primarily in how they explore phase space 
and thus compute thermodynamic properties such as the density of states $\Omega(E)$. The accuracy of any MC simulation depends on the details of how the system is `explored'. To reduce the error 
in simulation one can in principle simply continue to allow the simulation more time to explore 
phase space.

The motivation for new research into developing MC methods stems from the desire to solve ever-increasingly complex physical problems in a faster simulation time. Many of the originally developed MC methods are unable to effectively explore the complex phase space of real (and even many test) systems. Flat-histogram methods (introduced in Section~\HR{fhm}) are able to explore all energies and temperatures in a `single' simulation. In addition, they are able to explore complex energy landscapes without getting `stuck' (either never converging or converging at such a slow rate as to not practically converge). Many of these flat-histogram methods also have proven convergence on the statistical error thus guaranteeing convergence.
In the following sections, we will discuss a number of MC methods including importance sampling canonical methods, and flat-histogram methods leading up to our method SAD introduced in this work discussed in Section~\HR{fhm}.

\subsection{Random Number Generation}\label{rng}

Since Monte Carlo moves are inherently stochastic in nature, it is imperative that any useful random number generator be fast and ideally non-deterministic. In practice, pseudo-random number generators (PRNGs) are used in order to generate the fast random streams although they are deterministic. In order for a PRNG to be acceptable for use in Monte Carlo simulation it must meet certain criteria~\cite{sawilowsky2003you}. First, the PRNG must have a long `period' before repeating values. Poor PRNGs (that is ones with short periods) can lead to systematic error in Monte Carlo simulations~\cite{ferrenberg1992monte, plascak2002cluster, ossola2004systematic}.
Second, the PRNG must actually be able to pass certain statistical tests such as implemented in the TestU01 library~\cite{l2007testu01}.

There are a variety of PRNGs each with benefits and few without some form of drawback. We detail some of the well known PRNGs here. Linear congruential generators (LCGs) have a very fast time performance and are straightforward to program (in terms of size and complexity). Unfortunately, LCGs have multiple issues regarding the generated random number statistical quality~\cite{coddington1994analysis, barry1996recommendations}. While the period of the PRNG can be made acceptable by using 64-bit integers, the statistical correlations remain. Other PRNGs are the belong to the shift-register family (Xorshift RNGs). Like the LCG family, the shift-register RNGs fail many statistical tests~\cite{panneton2005xorshift} and are therefore often considered unacceptable for use in Monte Carlo simulations.

In our works, we use the \textbf{Xoroshiro128+} RNG which is named after its logical operations (XOR $\xrightarrow{}$ rotate $\xrightarrow{}$ shift $\xrightarrow{}$ rotate). The RNG has ultra-fast speed 0.72 $\nicefrac{\text{ns}}{\text{64-bits}}$ and incredible statistical quality~\cite{blackman2018scrambled}.

\subsection{Canonical Monte Carlo}\label{cmc}
In this section, we introduce canonical Monte Carlo and how it relates to flat-histogram methods introduced in Section~\HR{fhm}. The first importance sampling method developed was the famous Metropolis method~\cite{metropolis1953equation}.  The method works by choosing an initial state for the system $\Gamma$. Next, a random site $\Gamma_i$ is chosen and a move is proposed that will result in an energy change $\Delta E$ if accepted. The acceptance of the move is determined by whether a random number chosen on a range from zero to one is less than a Boltzmann factor. The process of moves and random number generation is repeated until some termination condition is met. This canonical method performs calculations at a fixed temperature $T$ only. Re-weighting can extrapolate in the vicinity of the simulated temperature, but this approach is limited. Another limitation of canonical Monte Carlo is that for complex energy landscapes and/or large systems the simulation can undergo a phenomena known as `critical slowing down' as the simulation will never or almost never visit the ground state energy $E_{\min}$~\cite{berg2000introduction}. These limitations are what drive the need for methods that can avoid getting `stuck' and that can simulate all temperatures in a single simulation.

\subsection{Flat-Histogram Methods}\label{fhm}
The goal of flat-histogram methods (also called \emph{broad-histogram}
or \emph{multicanonical} methods) is to simulate each energy with
similar accuracy.  In this way, they can accurately determine the density of states
over a broad range of energies and thus determine the thermodynamic
quantities such as heat capacity or internal energy
over a broad range of temperatures.
Properties that require more information, such as a spatial
correlation function or a response function, can still be computed
for any temperature, provided statistics are collected for each
individual energy, which can then be reweighted for any
temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
errington2003direct}.
All the flat-histogram Monte Carlo methods begin with randomly chosen
`moves' which change the state of the system and must satisfy
detailed balance.  All of the tested algorithms differ in how they determine the
probability of accepting a move and in what additional statistics must
be collected in order to decide on that probability.
Flat-histogram methods calculate the density of states $D(E)$ for a discrete set
of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
trebst2004optimizing}. Therefore, energy binning becomes an important
consideration for systems with a continuum of possible energies.  Energy bins
are typically of uniform size for the entire energy
continuum~\cite{fasnacht2004adaptive}.

All of the closely related flat-histogram
methods rely on a weight function $w(E)$.  In these
algorithms, the probability of accepting a move is given by
\begin{equation}
	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
\end{equation}
which biases the simulation in favor of energies with low weights.  A set of
weights that are proportional to the density of states $D(E)$ of the system will
result in an entirely flat histogram, which means that the weights should
converge to being proportional to the density of states.  The natural logarithm
of the weight is typically stored, since the density of states will often vary
over a few hundred orders of magnitude. In the microcanonical ensemble, the
entropy is defined as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the
logarithm of the weight is an approximation of the entropy.

Each approach uses a random walk in energy space to estimate the density of
states.  The core of these approaches is to update the weights at each step of
the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma_t
\end{equation}
where $t$ is the number of the current move, $\gamma_t$ is an update factor
that varies over the simulation, and $E$ is the current energy.  This update
causes the random walk to avoid energies that have been frequently sampled,
enabling a rapid exploration of energy space. This approach violates detailed
balance, due to the acceptance probabilities changing with each move, but the
severity of this violation decreases as we decrease $\gamma_t$.  The four
methods differ primarily in how they schedule the decrease of $\gamma_t$.

In this work, we develop a new method called Stochastic Approximation with Dynamic update factor (SAD) method which is a variant of the SAMC Algorithm that attempts to dynamically choose
the modification factor rather than relying on system dependent
parameters such as $t_0$ or $\gamma_0$.  There is an immediate
advantage of such an algorithm where parameters are chosen independent
of system size or type. Our
proposed method SAD requires the user to input $T_\text{min}$, the
lowest temperature of interest, which is an immediate disadvantage of
the method. However, identifying a minimum temperature of interest
$T_\text{min}$ may be easier for a user than determining in advance
an energy range of interest or the unphysical parameter $t_0$ such as other
flat-histogram methods must do.

\section{Test System Details}
We introduce a number of systems that are suitable for benchmarking the convergence properties of various flat-histogram MC methods. Testing on a variety of systems is necessary for ensuring the robustness of a given MC method. The square-well fluid is chosen as a test-bed due to the fact that convergence properties of flat-histogram weight-based MC methods have never been studied on this system. Secondly, the MC method will have to simulate the potentially complex liquid-vapor phase coexistence of the system. The 2D Ising model is a simple system with a second order phase transition. In addition, the majority of MC methods have been tested on the Ising model so the speed and accuracy with which a given MC method can converge can often be directly compared with the literature. A Lennard-Jones cluster is a heavily studied system that can be simple or incredibly complex for a MC method to accurately solve. In order to solve for solid-solid transitions for certain cluster configurations, it is necessary to explore incredibly low reduced temperatures. This can make finding the global minima configuration extremely difficult for MC methods in general.

\subsection{Square-Well Fluid}
The square-well fluid is a system of particles whose interactions are governed by a square-well potential~\cite{singh2003surface, barker2004perturbationSW}.  The square-well potential is an ideal test-bed for Monte Carlo methods as it is a simple model for a liquid, which includes both attractive and repulsive interactions~\cite{barker1967-SW-perturbation, vega1992phase}. The potential $U(\textbf{r})$ for such a system is given by
\begin{equation}
 U(\textbf{r})=\begin{cases} \infty &
 \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
 \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
 \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
\end{equation}
where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the reduced range of the potential well, and $\epsilon$ is its depth.

\subsection{2D Ising Model}
The 2D Ising spin-lattice system is widely used as a test-bed when benchmarking or comparing Monte-Carlo methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The \nth{2} order phase transition behavior and the ability to directly calculate the exact solution for finite lattices~\cite{beale1996exact} make the system sufficiently interesting for such theoretical comparisons.
The Hamiltonian for the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new} cab be written as the following:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the absence of a magnetic field ($h =
0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over nearest neighbor spin sites.

\subsection{Lennard-Jones Clusters}
Lennard-Jones cluster systems are widely used as a challenging test-bed when comparing Monte-Carlo methods~\cite{doye1998thermodynamics, poulain2006performances, martiniani2014superposition}. The potential energy for interacting atomic particles can be written as follows:
\begin{align}
U = 4\epsilon \left[\left( \sigma/r \right)^{12} - \left( \sigma/r \right)^{6}\right]
\end{align}
A number of challenges exit when trying to accurately simulate the density of states of a cluster system. Narrow potential energy funnels~\cite{mandelshtam2006multiple, wales1997global} and low temperature solid-solid transitions~\cite{neirotti2000phase, calvo2000phase, calvo2000entropic} make predicting thermodynamic properties such as heat capacity an extremely difficult process. In addition, since flat-histogram methods explore all energies and all temperatures it is possible that for high temperatures a large cluster volume is possible (which could lead to boil-off of the particles). A `constraining' sphere with radius $R_c$ is used to prevent this but it is usually empirically determined further complicating the simulation process.

@@

We have provided the background necessary for understanding the various flat-histogram methods and the various applied test systems. In the following chapters, we detail our manuscripts and the method SAD applied to the various physical test systems. We test SAD against a variety of other weight-based flat-histogram methods on the square-well fluid, Lennard-Jones clusters, and the 2D Ising model. Each of these systems provides unique challenges to our newly developed method SAD and convergence robustness is determined best by the application to multiple systems.
